Abstract Data Type Inference

### 2015-03-10 algebra.tex
Added references to nature.
Consistent Section 2.

Two histories $h_1$ and $h_2$ are related by $\to_x$, for $x = \mathrm{o},
\mathrm{c}, \mathrm{p}$, when $h_2$ is obtained from $h_1$ by:
\begin{itemize}

  \item removing an order constraint (o),

  \item making a completed operation pending (c), or

  \item adding a pending operation (p).

\end{itemize}
A set of histories $H$ is \emph{closed} under a relation $\to$ when $h_2 \in H$
whenever $h_1 \to h_2$ and $h_1 \in H$. A fundamental property of concurrent
object implementations is that their histories are closed under weakening via
less ordering, fewer operations completed, and additional pending
operations~\cite{conf/popl/BouajjaniEEH15}.

  $H(\mathcal{I})$ is closed under $\to_\mathrm{o}$, $\to_\mathrm{p}$, 
  $\to_\mathrm{c}$.

@@

\end{itemize}
We say an ADT closed under $\to_\mathrm{m}$ and $\to_\mathrm{d}$ is
\emph{normal}. Defining the relation $\succeq$ as the reflexive and transitive
closure of the above relations
\begin{align*}
  \succeq = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c} \cup 
  \to_\mathrm{m} \cup \to_\mathrm{d})^\ast,
\end{align*}
closure under $\succeq$ follows immediately.

@@

For all $i \in \mathbb{N}$, let
\begin{align*}
  B_i &= \set{ h \in B : \_ \le i } \\
  B_i' &= \set{ h' \in B :
    \exists h \in B_i.\ h \preceq h' \text{ and } \_ \le i+1
  }
\end{align*}
be, respectively, the histories of $B$ with at most $i$ matches, and those
derived from $B_i$ with at most $i\!+\!1$ matches. We say that $A$ is
\emph{predictable} if $B_i = B$ whenever $B_i' = B_{i+1}$.

>>>>

TODO MOTIVATE THIS SECTION

@@

\end{itemize}
We say an ADT closed under $\to_\mathrm{m}$ and $\to_\mathrm{d}$ is
\emph{normal}. In Section~\ref{sec:nature} we demonstrate that
naturally-occurring ADTs are normal. Defining the relation $\succeq$ as the
reflexive and transitive closure of the above relations
\begin{align*}
  \succeq = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c} \cup 
  \to_\mathrm{m} \cup \to_\mathrm{d})^\ast,
\end{align*}
closure under $\succeq$ follows immediately.

@@

For all $i \in \mathbb{N}$, let
\begin{align*}
  B_i &= \set{ h \in B : \_ \le i } \\
  B_i' &= \set{ h' \in B :
    \exists h \in B_i.\ h \preceq h' \text{ and } \_ \le i+1
  }
\end{align*}
be, respectively, the histories of $B$ with at most $i$ matches, and those
derived from $B_i$ with at most $i\!+\!1$ matches. We say that $A$ is
\emph{predictable} if $B_i = B$ whenever $B_i' = B_{i+1}$. In 
Section~\ref{sec:nature} we demonstrate that naturally-occurring ADTs are
predictable.

### 2015-03-10 inference.tex
Consistent Section 2.

\end{itemize}
A \emph{matching scheme} $M$ associates to each execution $e$ with operations
$O$ a partial matching function $M(e): O \rightharpoonup O$.

@@

\end{itemize}
Non-completed operations are \emph{pending}, and are maximal in the
happens-before order. The operations of a \emph{sequential} history are totally
ordered. An operation $o \in \rng(m)$ is a \emph{match target}, and the set
$\set{ o' \in O : m(o') = o}$ of operations targeting $o$ is a \emph{match}. We
assume targets target themselves: $o \in \set{ o' \in O : m(o') = o}$.

@@

An \emph{implementation} $\mathcal{I}$ is a prefixed-closed set of executions.
An \emph{abstract data type (ADT)} $A$ is the set $H(\mathcal{I})$ of histories
of an implementation $\mathcal{I}$. We define \emph{observational refinement}
between two implementations, or between an implementation and an ADT, as
history-set inclusion. We say $\mathcal{I}_1$ \emph{refines} $\mathcal{I}_2$
if{f} $H(\mathcal{I}_1) \subseteq H(\mathcal{I}_2)$, and $\mathcal{I}$
\emph{refines} $A$ if{f} $H(\mathcal{I}) \subseteq A$.

@@

TODO DEFINE KERNEL AND COMPLEMENT

A history formula $F$ \emph{represents} an ADT $A$ when:
\begin{itemize}

>>>>

\end{itemize}
An \emph{implementation} $\mathcal{I}$ is a prefix-closed set of executions
which is additionally closed under
\begin{itemize}

  \item appending call actions (of fresh operations),

  \item permuting call actions backward in time, and

  \item permuting return actions forward in time.

\end{itemize}
These conditions capture the environment in which implementations execute:
calls are always enabled in their client programs, and thread schedulers may
induce arbitrary delay between implementation code and the associated call and
return actions~\cite{conf/popl/BouajjaniEEH15}.

@@

\end{itemize}
Non-completed operations are \emph{pending}, and are maximal in the
happens-before order. The \emph{width} of $h$ is the maximal size of a
subset of $O$ for which each pair of operations is unordered. The \emph{width}
of a history set is the maximal width of its elements. Width-$1$ histories are
\emph{sequential}. An operation $o \in \rng(m)$ is a \emph{match target}, and
the set $\set{ o' \in O : m(o') = o}$ of operations targeting $o$ is a
\emph{match}. We assume targets target themselves: $o \in \set{ o' \in O :
m(o') = o}$.

Our abstraction from executions to histories relies correlating the operations
associated with the same values via matching. A \emph{matching scheme} $M$
associates to each execution $e$ with operations $O$ a partial matching
function $M(e): O \rightharpoonup O$.

@@

Two histories $h_1$ and $h_2$ are related by $\to_x$, for $x = \mathrm{o},
\mathrm{c}, \mathrm{p}$, when $h_2$ is obtained from $h_1$ by:
\begin{itemize}

  \item removing an order constraint (o),

  \item making a completed operation pending (c), or

  \item adding a pending operation (p).

\end{itemize}
A set of histories $H$ is \emph{closed} under a relation $\to$ when $h_2 \in H$
whenever $h_1 \to h_2$ and $h_1 \in H$. A fundamental property of
implementations is that their histories are closed under weakening via less
ordering, fewer operations completed, and additional pending
operations~\cite{conf/popl/BouajjaniEEH15}.

  $H(\mathcal{I})$ is closed under $\to_\mathrm{o}$, $\to_\mathrm{p}$, 
  $\to_\mathrm{c}$.

An \emph{abstract data type (ADT)} $A$ is the set $H(\mathcal{I})$ of histories
of an implementation $\mathcal{I}$. A \emph{kernel} of an ADT $A$ is a minimal
set $H$ of histories such that
\begin{align*}
  A = \set{ h' : \exists h \in H.\ h \to^\ast h' }
\end{align*}
where $\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$. In
Section~\ref{sec:nature} we demonstrate that naturally-occurring ADTs have
unique kernels. For the remainder of this work, we assume that ADTs have unique
kernels, and denote \emph{the kernel} of $A$ by $\ker A$. We say that $A$ is
\emph{atomic} when the width of $\ker A$ is $1$.

We define \emph{observational refinement} between two implementations, or
between an implementation and an ADT, as history-set inclusion. We say
$\mathcal{I}_1$ \emph{refines} $\mathcal{I}_2$ if{f} $H(\mathcal{I}_1)
\subseteq H(\mathcal{I}_2)$, and $\mathcal{I}$ \emph{refines} $A$ if{f}
$H(\mathcal{I}) \subseteq A$.

@@

The \emph{complement} of an ADT $A$ with width $k \in \mathbb{N}$ is the set of
histories of width at most $k$ which are excluded from $A$. A history formula
$F$ \emph{represents} an ADT $A$ when:
\begin{itemize}
